---
title: "PRL Gas Flux Analysis"
author: "Gavin McNicol"
date: "2026-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r}

install.packages("tidyverse")
library(tidyverse)
library(readxl)
library(openintro)
library(tidyr)
library(dplyr)
library(lubridate)
library(tidymodels)
library(data.table)
library(patchwork) 
library(hrbrthemes)
library(multilevelmod)
library(lme4)
library(nlme)
detach(package:lme4)
library(parsnip)
install.packages("ggpubr")
library(ggpubr)
library(ggsignif)
install.packages("ggtern")
library(ggtern)
library(ggplot2)
install.packages("gridExtra")
library(gridExtra)
install.packages("data2")
library(data2)
library(ggplot2)
install.packages("multcompView")
library(multcompView)
library(phyloseq)
```

get data

```{r}
data <- read_csv("francois-prl-fluxes.csv")
```

```{r}
data
```

summarize fluxes and output as csv

```{r}
data %>% 
  group_by(date_collected, treatment, gas) %>% 
  summarize(mean_flux_umol = mean(flux_umol_m2_s1, na.rm = T),
            mean_flux_nmol = mean(flux_nmol_m2_s1, na.rm = T)) %>% 
  arrange(gas) %>% 
  write_csv("francois-prl-fluxes-summary.csv")
```

linear regression and/or anova can go here....

```{r juniors-code}
lm(y ~ x)
```

# JUnior continues on Gavin data

## Average CO2 flux by treatment

```{r}
data %>% 
  filter(gas=="co2") %>% 
  group_by(treatment) %>% #date_collected) %>% 
  summarise(mean_flux= mean(flux_umol_m2_s1, rm.na= TRUE))
```


# Stat analysis

```{r}
anova_co2 <- aov(flux_umol_m2_s1 ~ treatment, data = data)
summary(anova_co2)
```

```{r}
#co2_data<- 
data %>% 
  filter(gas=="co2")
```


# Import PRL flux data for 10-02-25 and 10-13-25

```{r}
flux_data <- read.csv("/Users/Junior/Documents/PRL_experiment_2025/all_date-flux-summary-PRL.csv")%>% mutate(date_collected = as.Date(date_collected, format = "%m/%d/%Y"))
```


```{r}
flux_data
```

# CO2 flux vs treatments

```{r}
flux_data %>% 
  filter(gas=="co2") %>% 
  group_by(treatment) %>% #date_collected) %>% 
  summarise(mean_flux= mean(co2_flux, rm.na= TRUE))
```


# Test ANOVA for significance in co2 flux between treatments

```{r}
anova_co2 <- aov(co2_flux ~ treatment, data = flux_data)
summary(anova_co2)
```

# Ch4 flux vs treatments

```{r}
flux_data %>% 
  filter(gas=="ch4") %>% 
  group_by(treatment) %>% # date_collected) %>% 
  summarise(mean_flux= mean(ch4_flux, rm.na= TRUE))
```

# Test ANOVA for significance in ch4 flux between treatments

```{r}
anova_ch4 <- aov(ch4_flux ~ treatment, data = flux_data)
summary(anova_ch4)
```

# n2o flux vs treatments

```{r}
flux_data %>% 
  filter(gas=="n2o") %>% 
  group_by(treatment) %>% #date_collected) %>% 
  summarise(mean_flux= mean(n2o_flux, rm.na= TRUE))
```

# Test ANOVA for significance in no2 flux between treatments

```{r}
anova_n2o <- aov(n2o_flux ~ treatment, data = flux_data)
summary(anova_n2o)
```

#########
